\documentclass{letter}

\signature{Alex Groce}
%\address{21 Bridge Street \\ Smallville \\ Dunwich DU3 4WE}
\begin{document}

\begin{letter}{}
\opening{Dear reviewers:}

I submit this revision to my Onward! Essay, which I believe addresses the concerns noted by reviewers.

The proposed changes from the meta-review are:

\begin{enumerate}

\item {\bf Page 1 "(Figure 1))"} The figure mention looks correct now, and clicking in the PDF works.

\item {\bf Epsilon says, "The Turing Award in 2007 was given for this..."}  The receipients (including my late advisor!) are now noted.

\item {\bf  References in the paper should be more explicit for the positions that the individual pupils defend.}  The paper now notes that connection between pupils and positions is less definite than this; however, the references were very useful, and have all been incorporated as ``pivots'' around which the students might be rotating.  I did not include Lehman, because I think the connection between what I/the pupils meant and Lehman's notions was too weak for this to be a workable inclusion.  The Dijkstra reference forms the basis of a substantial new section at the end of the essay, and I very much thank reviewers for directing my attention to it!

\item {\bf Adding to section 2, when discussing the idea of differential testing: the oracle problem is important, as the first is a solution to the latter (and maybe what the students tried to do with the linear search idea).}  I hope this, with the commentary (see below) is now clearer.

\item {\bf Regarding the last section: it wrapped the techniques the students talked about.}  I incorporated the core of the postscript, plus a substantial expansion on each part of it, as a running commentary that occasionally ``interrupts'' the flow.  I tried parallel columns but found this was actually too non-linear to be very readable, but stopping the dialogue to discuss what's (recently) been said seemed to work well.

\item {\bf A bit more context regarding today's software which is a lot around LLM.}  This is included in the running commentary, after the students mention LLMs, now.

\end{enumerate}

In general, the changes to the dialogue itself are minimal; reading the boxed {\bf COMMENTARY} sections should be sufficient to check for the requested changes.

  I very much enjoyed reading the Petricek essay, and have noted the serendipitous similarity of approach and broad topics in the new, reduced, Postscript.

\closing{Sincerely,}



\end{letter}
\end{document}